---
title: "Hawaiian Basalts"
author: "Julie M. Coulombe"
date: "11/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
```

## Alkali-Basalt vs Tholeiite

The dataset was pulled from <https://search.earthchem.org/> with Feature Name = OceanIsland:Hawaii and Sample Types = igenous:volcanicmafic:alkali.basalt, tholeiite, trachyte, trachytebasalt


```{r mydata}
alkali_thol <- select(mydata, "SAMPLE.ID", "METHOD":"ZR") %>%
  filter(ROCK.NAME == "ALKALI BASALT" | ROCK.NAME == "THOLEIITE")
cnt_alk <- table(alkali_thol$ROCK.NAME)
cnt_alk
```

This data set did not include enough information to analyze - due to only having
1 alkali basalt.  When trying to filter for only sio2, that data was not available for any of the observations.  At this point I went back to the original database and pulled a new dataset.

```{r mydata_2}
alk_thol <- select(mydata_2, "SAMPLE.ID", "METHOD":"ZR") %>%  
  filter(ROCK.NAME == "ALKALI BASALT" | ROCK.NAME == "THOLEIITE")
cnt_rocks <- table(alk_thol$ROCK.NAME)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
