---
title: "Hawaiian Basalts"
author: "Julie M. Coulombe"
date: "11/30/2019"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
mydata<-read.csv("C:/Users/labry/Documents/R/final_project_490/data/elmnts_data.csv")
mydata_2 <- read.csv("C:/Users/labry/Documents/R/final_project_490/data/earthchem_dataset2.csv")
```

## Alkali-Basalt vs Tholeiite

The dataset was pulled from <https://search.earthchem.org/> with Feature Name = OceanIsland:Hawaii and Sample Types = igenous:volcanicmafic:alkali.basalt, tholeiite, trachyte, trachytebasalt


```{r mydata}
alkali_thol <- select(mydata, "SAMPLE.ID", "METHOD":"ZR") %>%
  filter(ROCK.NAME == "ALKALI BASALT" | ROCK.NAME == "THOLEIITE")
cnt_alk <- table(alkali_thol$ROCK.NAME)
cnt_alk
```

This data set did not include enough information to analyze - due to only having
1 alkali basalt.  When trying to filter for only sio2, that data was not available for any of the observations.  At this point I went back to the original database and pulled a new dataset.

```{r mydata_2}
alk_thol2 <- select(mydata_2, "SAMPLE.ID", "METHOD":"ZR") %>%  
  filter(ROCK.NAME == "ALKALI BASALT" | ROCK.NAME == "THOLEIITE")
cnt_rocks2 <- table(alk_thol2$ROCK.NAME)
cnt_rocks2
```

This is great - lots of alkali and tholeiite to pull from
## Let's see if we can plot the elements for alkali and basalt

```{r Plot of SIO2 vs Na2O & K2O}
alk_thol_plot <- ggplot(alk_thol2, aes(x = SIO2, y = NA2O + K2O)) +
  geom_point()
alk_thol_plot
```

This plot looks a little crazy - determined that data pull from database did not filter for Hawaii samples only.  Let's try to do this in R

```{r filter for Hawaii}
alk_thol3 <- select(mydata_2, "SAMPLE.ID":"TITLE", "METHOD":"ZR") %>% 
  filter(grepl("HAWAII", mydata_2$TITLE)) %>%
  filter(ROCK.NAME == "ALKALI BASALT" | ROCK.NAME == "THOLEIITE")
cnt_rocks3 <- table(alk_thol3$ROCK.NAME)
cnt_rocks3
```

so after all that work, it turns out that when filtered to "hawaii", I am left
with no alkali basalt samples to compare against

So let's look at Hawaii and some major element data, and compare Mg to Si

```{r filter for SIO2 data}
sio2_sort <- select(mydata_2, "SAMPLE.ID":"TITLE", "METHOD":"ZR") %>% 
  filter(mydata_2$SIO2 != "NA") %>%  
  filter(grepl("HAWAII", TITLE))
```

and the plot for that filter selection
```{r Plot SIO2 vs MgO}
sio2_plot2 <- ggplot(sio2_sort, aes(x = SIO2, y = MGO)) + 
  geom_point() +   
  geom_smooth() +  
  ggtitle("SiO2 compared to MgO") +  
  xlab("SiO2") + 
  ylab("MgO")
sio2_plot2
```